---
title: "Regression Models Course Project"
output: 
  pdf_document: 
    fig_crop: no
---

Author: **Fred Zhou**

In this document, we will try to answer the following questions:

- Q1: "Is an automatic or manual transmission better for MPG"

- Q2: "Quantify the MPG difference between automatic and manual transmissions"

By default, we assume that for the `mpg`, the lower the value the better. 

(For `am`, `0` for automatic transmission, `1` for manual transmission.)


###Q1. Is an automatic or manual transmission better for MPG
To answer this question, we assume that the all the variables in the population follow normal distribution. Thus we first use Student's T test to address whehter there's difference in these two groups


####Student's T-test between AUTOMATIC and MANUAL (alpha=0.05)
```{r T-test, echo=TRUE,cache=TRUE}
test_mpg=t.test(mtcars$mpg[mtcars$am==1],mtcars$mpg[mtcars$am==0])
print(test_mpg)
print(paste(
'The P-value for the T-test between AUTOMATIC and MANUAL transmissions for the mpg is ',
round(test_mpg$p.value,digits = 4),sep=''))
print(paste(
'Mean value for the mpg with AUTOMATIC transmissions:',
round(test_mpg$estimate[1],digits = 2),sep=''))
print(paste(
'Mean value for the mpg with MANUAL transmissions:'
,round(test_mpg$estimate[2],digits = 2),sep=''))
```

Thus we could address that indeed the types of transmission will affect the `mpg`, and on average `AUTOMATIC` will bear a *higher consumption of fuel* against the `MANUAL` transmission, and the average difference is around *7.24* miles per Gallon used.

###Q2. Quantify the MPG difference between automatic and manual transmissions
####Correlation analysis winthin all variables against the mpg
```{r covariant, echo=TRUE,cache=TRUE}
sort(abs(cor(mtcars)[1,]))
```
We already get the hint that the `AUTOMATIC/MANUAL` have impacts on the fuel consumption, thus from the correlation analsis we could guess that any variant with a higher correlation value against `AUTOMATIC/MANUAL` may contribute to the fuel consumption. including:

1.`vs` - V/S

2.`drat` - Rear axle ratio

3.`hp` - Gross horsepower

4.`disp` - Displacement (cu.in.)

5.`cyl` - Number of cylinders

6.`wt` - Weight (1000 lbs)

Thus, we could guess that it's reasonable to include any variable into the linear regressions. We could make a most general form of regression, then add in more variants to further optimize our model.


#### General model
We only take the `am` as variables to do the linear regression first:
```{r regression_1, echo=TRUE,cache=TRUE}
fit_1 <- lm(mpg~am, data = mtcars)
summary(fit_1)
```
Based on the stat data we could address:

- On average, AUTOMATIC car have 17.15 MPG and MANUAL transmission cars have 7.25 MPG more

- The R^2 value is only 0.36, which means that our current model only explains 36% of the variance


#### Multivariate model - adapted selection of variants
```{r regression_2, echo=TRUE,cache=TRUE}
fit_2 = step(lm(data = mtcars, mpg ~ .),trace=0,steps=50000)
summary(fit_2)
```

###Summary
This model explains 84% of the variance in miles per gallon (`mpg`), which is accaptable for the predicion of `mpg` with new data.
Based on the multivariate model we could address:

- `MANUAL` is  beneficial for the fuel saving, after model adjusting the value comes to be *2.936* miles per gallon.

- `wt` affect huge against the `mpg`, which is appearant since more load will eventually consume more fuel.

###APPENDIX 

####Visualize the data between AUTOMATIC and MANUAL
```{r Visualize, echo=FALSE,cache=TRUE,fig.align='left'}
boxplot(mpg ~ am,data=mtcars,
        ylab='mpg (Miles per Gallon)', 
        xlab='Transmission',
        names=c('automatic','manual'),
        main='Transmission vs mpg')
```

####Comparision of general and multivariate model

1.ANOVA

```{r ANOVA, echo=TRUE,cache=TRUE}
anova(fit_2, fit_1)
```

2.Residual diagnostics for multivariate model

```{r residue, echo=TRUE,cache=TRUE}
par(mfrow = c(2,2))
plot(fit_2)
```