---
title: "Statistical_inference_2"
output: pdf_document
---

Author: *Fred Zhou*

Using `ToothGrowth` data and compare odontoblasts growth by `supp` and `dose`

##Basic summary

```{r}
#Import data
data(ToothGrowth)
#Visualize data
head(ToothGrowth)
#change dose to factor
ToothGrowth$dose=factor(ToothGrowth$dose)
#Basic summary
summary(ToothGrowth)
#Plot visulaization
boxplot(len ~ supp + dose, data=ToothGrowth,ylab='odontoblasts len',main='odontoblasts len vs supplement and dosage ')
```

### Guess
- **Vitamine C could facilite the growth of odontoblasts. **

- **Low and medium dose (0.5 and 1 mg/day) of VC supplied with orange juice seems to favor the odontoblasts growth of guinea pigs compared with ascorbic acid.**

- **For high dose (2 mg/day), the effect of orange juice and ascorbic acid seems to be same.**


##1. Effect of supplement and dosage

Assume that variance in each condition are the same. The Two-way ANOVA is adopted to see the effect of both supply and dosage.


The NULL hypothesis `H0`: 

1.The population means of `len` in `OJ` and `VC` (`supp`) are the same.

2.The population means of `len` in `0.5`, `1` and `2` (`dose`) are the same.

3.There is no interaction between the two factors (`supp` and `dose`).

```{r}
#Two-way ANOVA
two_way=aov(len ~ supp*dose,data=ToothGrowth)
summary(two_way)
```

Result reject all items in NULL hypothesis `H0` in the `95%` confidient intervals.
Which indicate that:

1. The `supply` affects the odontoblasts growth (`P-value = 0.000231`)

2. The `dose` affects the odontoblasts growth (`P-value < 2e-16`)

3. There is interaction between `OJ` and `VC` (`P-value = 0.021860`), which means that the supply may affect the Vitamine C functioning on the pig.

We will further test effect within each group using the `Two-way ANOVA Tukey post-hoc test` to test the mean value between differnt groups.



##2. Effect on the odontoblasts growth

All the NULL hypothesis `H0` is the mean value is same.

```{r}
#Two-way ANOVA
TukeyHSD(two_way)
```


- In the `$supp` test, the NULL hypothesis `H0` is rejected (`adj P-value < 0.0002312`), indicating the effects for the `VC` and `OJ` is statistically different, with mean value of odontoblasts length 3.7 longer in OJ treatment.

- In the `$dose` test, the NULL hypothesis `H0` is rejected in all the dosase comparement (`adj P-value = 2.7e-06 or 0`), indicating the positive effect of Vitamine C on the growth of odontoblasts, and the effect is most obvious between dosage `1` and `0.5`

- In the `$supp:dose` test, only in following gropus no significant difference is found (95% confidence interval):

1. `VC:1-OJ:0.5` (`adj P-value = 0.2640208`)

2. `OJ:2-OJ:1` (`adj P-value = 0.3187361`)

3. `VC:2-OJ:1` (`adj P-value = 0.2936430`)

4. `VC:2-OJ:2` (`adj P-value = 1`)

The effect of Vitamine C tend to be satured for the growth of odontoblasts at dosage of 2 mg/day. All the guesses have been proved.