---
title: "National Weather Service Storm Data"
output: html_document
---
### Author: *Fred Zhou*

## Question needs to answer:
- **Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health?**

  The sum of FATALITIES and INJURIES will be taken against each EVTYPE by function `aggregrate`, and the top 5 EVTYPE with values will be given by plot (ggplot2).

- **Across the United States, which types of events have the greatest economic consequences?**

  The sum of PROPDMG and CROPDMG will be determined by their EXPressions and sum against each EVTYPE by function `aggregrate`, and the top 5 EVTYPE with values will be given by table.
  
## Data Processing

### Loading and preprocessing the data

```{r read_data,echo=T,cache.vars=data,cache=TRUE}
setwd('/Users/fred/Desktop/datasciencecoursera/R/Reproducible_Research/')
data=read.table('/Users/fred/Downloads/repdata-data-StormData.csv',sep=',',stringsAsFactors=F,header=T,nrows=-1)
```

### Pre-processing for Question 1

```{r Harmful_EVTYPE,echo=T,results='asis',warning=F,cache=TRUE,warning=F,fig.width=12, fig.height=6}
library(xtable)
library(ggplot2)
Harmful_EVTYPE=aggregate(data[,23:24],by=list(data[,8]),function(x) sum(x,na.rm=T))
FATALITITES_EVTYPE=Harmful_EVTYPE[order(-Harmful_EVTYPE$FATALITIES),][1:5,]
FATALITITES_EVTYPE$Class=rep('FATALITITES',5)
INJURIES_EVTYPE=Harmful_EVTYPE[order(-Harmful_EVTYPE$INJURIES),][1:5,]
INJURIES_EVTYPE$Class=rep('INJURIES',5)

Merge=rbind(FATALITITES_EVTYPE,INJURIES_EVTYPE)
names(Merge)[1]='EVTYPE'

```

### Pre-processing for Question 2

```{r Money_damage,echo=T,results='asis',warning=F,cache=TRUE,warning=F,fig.width=12, fig.height=6}
library(xtable)
library(ggplot2)

data$PROPDMGEXP_new[grepl("[0-9]",data$PROPDMGEXP)]=10^as.numeric(data$PROPDMGEXP[grepl("[0-9]",data$PROPDMGEXP)])
data$PROPDMGEXP_new[is.na(data$PROPDMGEXP) | data$PROPDMGEXP=='-' |data$PROPDMGEXP =='?']=0
data$PROPDMGEXP_new[grepl('-|?',data$PROPDMGEXP)]=0
data$PROPDMGEXP_new[data$PROPDMGEXP=='B' |data$PROPDMGEXP=='b']=10^9
data$PROPDMGEXP_new[data$PROPDMGEXP=='H' |data$PROPDMGEXP=='h']=10^2
data$PROPDMGEXP_new[data$PROPDMGEXP=='K' |data$PROPDMGEXP=='k']=10^3
data$PROPDMGEXP_new[data$PROPDMGEXP=='M' |data$PROPDMGEXP=='m']=10^6

data$CROPDMGEXP_new[grepl("[0-9]",data$CROPDMGEXP)]=10^as.numeric(data$CROPDMGEXP[grepl("[0-9]",data$CROPDMGEXP)])
data$CROPDMGEXP_new[is.na(data$CROPDMGEXP) | data$CROPDMGEXP=='-' |data$CROPDMGEXP =='?']=0
data$CROPDMGEXP_new[grepl('-|?',data$CROPDMGEXP)]=0
data$CROPDMGEXP_new[data$CROPDMGEXP=='B' |data$CROPDMGEXP=='b']=10^9
data$CROPDMGEXP_new[data$CROPDMGEXP=='H' |data$CROPDMGEXP=='h']=10^2
data$CROPDMGEXP_new[data$CROPDMGEXP=='K' |data$CROPDMGEXP=='k']=10^3
data$CROPDMGEXP_new[data$CROPDMGEXP=='M' |data$CROPDMGEXP=='m']=10^6


data$PROP_total=data$PROPDMG*data$PROPDMGEXP_new
data$CROP_total=data$CROPDMG*data$CROPDMGEXP_new
data$Money_total=data$PROP_total+data$CROP_total

Money_EVTYPE=aggregate(data$Money_total,by=list(data$EVTYPE),function(x) sum(x,na.rm=T))

Money_EVTYPE[,2]=round(Money_EVTYPE[,2]/10^9,digits = 2)
names(Money_EVTYPE)=c('EVTYPE','Money_total (Billions)')
Money_EVTYPE=Money_EVTYPE[order(-Money_EVTYPE$Money_total),]

```

##Results

### Question 1

```{r Q1_plot,echo=T,results='asis',warning=F,cache=TRUE,warning=F,fig.width=12, fig.height=6}
qplot(INJURIES,FATALITIES,data=Merge,facets=Class~EVTYPE, 
   main="Top 5 Event for FATALITIES and INJURIES",size=I(3),shape=Class,
   xlab="", ylab="Sum of affected people number ",ylim=c(100,10000))
```

### Question 2

```{r Q2_Table,echo=T,results='asis',warning=F,cache=TRUE,warning=F,fig.width=12, fig.height=6}
library(xtable)
library(ggplot2)

print(xtable(Money_EVTYPE[1:5,], digits=2),type='html')
```