---
title: "Practical Machine Learning - Course project"
output: html_document
---
**Author: Fred ZHOU**

In this work, we will try to apply the machine learning algorithm to classify the individual behavior based on data collected from the mobile device.

Data collected from the following publication:

*Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. *

## 0. Loading Libraries

```{r loading packages,echo=T,cache=TRUE,warning=FALSE,}
library(caret)
library(randomForest)
library(rpart) 
library(rattle)
set.seed(12345678)
```

## 1. Import data

```{r loading data,echo=T,cache=TRUE,cache=TRUE}
training=read.csv('./pml-training.csv',stringsAsFactors = T,header=T)
training_y=training$classe
testing=read.csv('./pml-testing.csv',stringsAsFactors = T,header=T)
```

## 2. Clean data 

### View dim of data

```{r clean data,echo=T,cache=TRUE}
dim(training)
```

### Cleaning NearZeroVariance Variables

```{r NearZeroVariance,echo=T,cache=TRUE}
NZV <- nearZeroVar(training, saveMetrics=TRUE)
Col_id=NZV[,4]
training=training[,Col_id]
```

### Removing ID column

```{r Removing ID,echo=T,cache=TRUE}
training=training[,-1]
```

### Clean variables with Too many NA (NA ratio < 0.6)

```{r Clean NA,echo=T,cache=TRUE}
Count_NA=apply(training,2,function(x) sum(is.na(x)))/dim(training)[1]
training=training[,Count_NA<0.6]
dim(training)
training$classe=training_y
testing=testing[,names(testing) %in% names(training)] 
```

## 3. Training data with different model.

### Split training into training and crossvalidation

```{r crossvalidation,echo=T,cache=TRUE}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
tmp=training
training <- tmp[inTrain, ]
crossvalid <- tmp[-inTrain, ]
dim(training)
dim(crossvalid)
```

### Train with Random forest

```{r Random forest,echo=T,cache=TRUE}
fit_rf=train(classe~ ., method='rf',data=training)
pred_training_rf=predict(fit_rf,training)
pred_cross_rf=predict(fit_rf,crossvalid)
accuracy(pred_training_rf, training$classe)
accuracy(pred_cross_rf, crossvalid$classe)
```

### Train with Generalized Boosted Regression Models

```{r gbm,echo=T,cache=TRUE}
fit_gbm=train(classe~ ., method='gbm',data=training)
pred_training_gbm=predict(fit_gbm,training)
pred_cross_gbm=predict(fit_gbm,crossvalid)
accuracy(pred_training_gbm, training$classe)
accuracy(pred_cross_gbm, crossvalid$classe)
```

### Train with Linear Discriminant Analysis

```{r lda,echo=T,cache=TRUE}
fit_lda=train(classe~ ., method='lda',data=training)
pred_training_lda=predict(fit_lda,training)
pred_cross_lda=predict(fit_lda,crossvalid)
accuracy(pred_training_lda, training$classe)
accuracy(pred_cross_lda, crossvalid$classe)
```


### Stack moultiple model with random forest

```{r Stack,echo=T,cache=TRUE}
pred_training=data.frame(pred_training_rf,pred_training_gbm,pred_training_lda,classe=training$diagnosis)
fit_stack=train(diagnosis~.,method='rf',data=pred_training)
pred_training_stack=predict(fit_stack,training)
pred_cross_stack=predict(fit_stack,crossvalid)
accuracy(pred_training_stack, training$classe)
accuracy(pred_cross_stack, crossvalid$classe)
```
